`r params$doc_title`

### 2022-12-20 scripts/MSc.Rmd
updated GSD simulations

---
title: "`r params$doc_title`"
author: | 
        | christoph.reich@med.uni-heidelberg.de 
        | 
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
params:
  date: !r Sys.Date()
  doc_title: "MSc Default Title"
output: 
  html_document:
    code_folding: hide
    theme: paper
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 4
    df_print: kable
    latex_engine: xelatex
bibliography: ../references/references.bib
---

@@

    # CONFOUNDER
    eaf.gsd,
    beta.gsd,
    se.gsd,
    p.value.gsd,
    # OUTCOME
    eaf.gbc,
    beta.gbc,
    se.gbc,
    # pvalue missing and needs to be added from GBC_dat
    # harmonize decision
    palindromic, ambiguous, mr_keep
  ) %>% 
  left_join(GBC_dat %>% select(SNP, p_value_out), by=c("SNP"="SNP")) %>% 
  rename(
    p.value.gbc = p_value_out
  ) %>% 
  relocate(
    p.value.gbc, .before = palindromic
  ) %>% 
  as_tibble()
```

@@

```{r MAF-Pval, warning=FALSE}
ggplot(data = my_data_harm, aes(x=eaf.bmi, y=-log10(p.value.bmi)))+
  geom_point(alpha=0.6)+
  geom_hline(yintercept=-log10(median(my_data$p_value_exp)), col="red", alpha=0.6)+
  geom_vline(xintercept=median(my_data$eaf_exp), col="red", alpha=0.6)+
  xlab("Effect allele frequency for exposure")+
  ylab("-log10(P-value) ")+
  labs(title = "MAF vs Pval",
       subtitle = "Define different genetic architectures",
       #col='', 
       #caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  theme_few()+
  scale_colour_few() ->fig02
fig02

@@

# Simulation Framework

@@

@@

@@

```{r funnel-plot}
p4 <- mr_funnel_plot(res_single)
p4[[1]] +
  ggthemes::theme_few()
```

@@

```{r MendelianRandomization-analysis, warning=FALSE}
# outsourced calculation to `perform_mr.R`
# data FOR estimates in `perform_mr.R` script and simulation in `simstudy.R` script
saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))

## load results
est <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est.rds", full.names = TRUE)))

>>>>

---
title: "`r params$doc_title`"
author: | 
        | christoph.reich@med.uni-heidelberg.de 
        | 
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
params:
  date: !r Sys.Date()
  doc_title: "MSc Default Title"
output: 
  html_document:
    code_folding: hide
    theme: paper
    toc: true
    number_sections: true
    toc_float: true
    toc_depth: 4
    df_print: kable
    latex_engine: xelatex
bibliography: ../references/references.bib
editor_options: 
  chunk_output_type: inline
---

@@

    # CONFOUNDER
    eaf.gsd,
    beta.gsd,
    se.gsd,
    p.value.gsd,
    # OUTCOME
    eaf.gbc,
    beta.gbc,
    se.gbc,
    # pvalue missing and needs to be added from GBC_dat
    # harmonize decision
    palindromic, ambiguous, mr_keep
  ) %>% 
  left_join(GBC_dat %>% select(SNP, p_value_out), by=c("SNP"="SNP")) %>% 
  rename(
    p.value.gbc = p_value_out
  ) %>% 
  relocate(
    p.value.gbc, .before = palindromic
  ) %>% 
  mutate(
    p.value.gsd = ifelse(p.value.gsd == "<.0001",  # convert from string
                         yes = "0.0001", 
                         no=p.value.gsd), 
    p.value.gsd = as.numeric(p.value.gsd)) %>% 
  as_tibble()
```

@@

```{r MAF-Pval, warning=FALSE}
ggplot(data = my_data_harm, aes(x=eaf.bmi, y=-log10(p.value.bmi)))+
  geom_point(alpha=0.6)+
  geom_hline(yintercept=-log10(median(my_data$p_value_exp)), col=ggthemes_data$few$colors$Dark[3,2][[1]], alpha=0.6)+
  geom_vline(xintercept=median(my_data$eaf_exp), col=ggthemes_data$few$colors$Dark[3,2][[1]], alpha=0.6)+
  xlab("Effect allele frequency for exposure")+
  ylab("-log10(P-value) ")+
  labs(title = "MAF vs Pval",
       subtitle = "Define different genetic architectures",
       #col='', 
       #caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  theme_few()+
  scale_colour_few() ->fig02
fig02

@@

# A *Gallstone Disease* Simulation Framework

```{r MendelianRandomization-analysis-GSD, warning=FALSE}
# outsourced calculation to `perform_mr.R`
# data FOR estimates in `perform_mr.R` script and simulation in `simstudy.R` script
saveRDS(my_data_harm, paste0("../output/Rdata/", Sys.Date(), "_my_data_harm.rds"))

## load results
est.bmi.gsd <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est_BMI_GSD.rds", full.names = TRUE)))

est.bmi.gsd %>% 
  mutate(OR=exp(estimate), CI.lower=exp(estimate-se.estimate),CI.upper=exp(estimate+se.estimate)) %>%  
  kable(digits=3, caption = "MR Results for GSD (outcome) using genetic variants as instrumental variables for BMI") %>% 
  kable_classic(full_width = FALSE, html_font = "Cambria")
```

### Scatter Plot

```{r scatter-MendelianRandomization-GSD, fig.width=10, fig.height=6, warning=FALSE, message=FALSE}
# create mr.object
mr.obj = MendelianRandomization::mr_input(
    # exposure
    bx = my_data_harm$beta.bmi, 
    bxse = my_data_harm$se.bmi, 
    # outcome
    by = my_data_harm$beta.gsd, 
    byse = my_data_harm$se.gsd, 
    snps = my_data_harm$SNP,
    exposure = "Body mass index",
    outcome = "Gallstone disease"
  )

p1.gsd <- mr_plot(mr.obj, error = TRUE, 
                  line = "ivw", 
                  orientate = FALSE, 
                  interactive = FALSE, 
                  labels = FALSE)

p1.gsd$layers[[1]]$aes_params$size <- 1
p1.gsd$layers[[1]]$aes_params$alpha <- 0.5

p1.gsd$layers[[4]]$aes_params$colour <- ggthemes_data$few$colors$Dark[1,2][[1]]
p1.gsd$layers[[5]]$aes_params$colour <- ggthemes_data$few$colors$Dark[1,2][[1]]
p1.gsd$layers[[4]]$aes_params$alpha <- 0.3
p1.gsd$layers[[5]]$aes_params$alpha <- 0.3

p1.gsd$layers[[6]]$aes_params$colour <- ggthemes_data$few$colors$Dark[2,2][[1]]

p1.gsd+
  ggthemes::theme_few() -> p1.gsd

p1.gsd

# SAVE
filename.mr.scatter.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrscatter_gsd.png")
ggsave(filename = filename.mr.scatter.gsd, plot = p1.gsd, 
         width = 10, height = 6, 
         units = "in"  # default
         )
```

### Forest Plot

```{r forestplot-MendelianRandomization-GSD, fig.width=5, fig.height=24, warning=FALSE}
p2.gsd <- mr_forest(
  mr.obj,
  alpha = 0.05,
  snp_estimates = TRUE,
  methods = "ivw",
  ordered = TRUE
) +
  ggthemes::theme_few()

#SAVE
filename.mr.forest.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrforest_gsd.png")
ggsave(filename = filename.mr.forest.gsd, plot = p2.gsd, 
       width = 5, 
       height = 24, 
         #units = "in"  # default
         )
```

### Funnel Plot

```{r funnel-plot-GSD, fig.height=3, fig.width=5}
p4.gsd <- mr_funnel(mr.obj, CI=TRUE)+
  ggthemes::theme_few()

p4.gsd$layers[[1]]$aes_params$size <- 1
p4.gsd$layers[[1]]$aes_params$alpha <- 0.8
# linerange layer
p4.gsd$layers[[2]]$aes_params$alpha <- 0.2
# xintercept layer - estimate
p4.gsd$layers[[4]]$aes_params$colour <- ggthemes_data$few$colors$Dark[9,2][[1]]
# delete linerange layer
p4.gsd$layers <- p4.gsd$layers[-2]

# plot
p4.gsd

#SAVE
filename.mr.funnel.gsd <- paste0("../output/plots/", Sys.Date(), "_figure_mrfunnel_gsd.png")
ggsave(filename = filename.mr.funnel.gsd, plot = p4.gsd, 
       width = 8, 
       height = 4, 
       units = "in"  # default
         )

# B *Gallbladder Cancer* Simulation Framework

@@

### Scatter Plot

@@

### Forest Plot

@@

### Funnel Plot

```{r funnel-plot, fig.height=3, fig.width=5}
p4 <- mr_funnel_plot(res_single)
p4[[1]] +
  ggthemes::theme_few() -> p4

p4

#SAVE
filename.mr.funnel.gbc <- paste0("../output/plots/", Sys.Date(), "_figure_mrfunnel_gbc.png")
ggsave(filename = filename.mr.funnel.gbc, plot = p4, 
       width = 8, 
       height = 4, 
       units = "in"  # default
         )
```

@@

```{r MendelianRandomization-analysis, warning=FALSE}
## load results
est <- readRDS(file = dplyr::last(list.files("../output/Rdata/", pattern = "_est.rds", full.names = TRUE)))
